---
title: "analysis_baseline_area"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(slam)
library(gurobi)
library(spatstat)
library(maptools)
library(rgdal)
library(terra)
library(rgeos)

```


```{r}
#close and rerun all of the rasters individually
#check all layers in the geomorph stacks to get rid of the lines
#remove any line rasters
#rerun geomorph and all analyses
#resave as tifs 
#save out as pngs
#add to report

```

## Scenario 1 - Baseline
The objective of this scenario was to identify important areas based on critical species data and distinct habitats and geomorphological features using a uniform cost for all planning units. For this run, we used area as the cost, set targets of 20% for all conservation features, locked in planning units that included existing MPAs, and set a boundary penalty to zero. 

### STEP 1: Assign planning units, cost layer, and conservation feature

#### Planning Unit
```{r}
###read in the planning unit 

pu <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_PU_grid.tif'))

plot(pu)

#get the boundary of the area
#make all values the same
r <- pu > -Inf
# convert to polygons (you need to have package 'rgeos' installed for this to work)
GA_polygon <- rasterToPolygons(r, dissolve=TRUE)
plot(GA_polygon)

writeOGR(GA_polygon, dsn = here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_area_outline.shp'), layer = 'poly', driver = "ESRI Shapefile")


```

#### Cost Layer
Cost layer is using area as the cost to establish a baseline scenario.
```{r}

### pull in the initial MZ EEZ raster to start with
GA_rast <-  raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_PU_grid.tif'))

### assign a uniform cost value of 0.5 to all cells since the area of the planning units are the same. 
### We assigned a value of 0.5 since we will transform cost layers in future simulations to be between 0-1 and this 0.5 is a middle value in this range 
values(GA_rast) <- 0.5

### mask the area raster to just the MZ EEZ, so all cells outside the EEZ are assigned NA AND rename it to cost_area
GA_area <- mask(GA_rast, pu)

### check it by plotting!
#plot(cost_area)

plot(GA_area)

### write raster
writeRaster(GA_area, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_PU_0.5.tif'), overwrite = TRUE)


GA_area_image <- as.im(GA_area)
plot(GA_area_image)


```

#### Conservation Features
This includes all species distributions and habitat rasters

Creating Species Stack
```{r}
### create file path to read in all 30 species tifs in 
species_path <- 'G:/group_project/future4fins/future4fins/set_up/final_rasters/conservation_features/iucn_species'
species_files <- list.files(species_path, full.names = TRUE)

### stack them 
species_stack <-  stack(species_files)

### plot to see how they look
plot(species_stack)

### write raster
writeRaster(species_stack, here('set_up/final_rasters/species_stack.tif'), overwrite = TRUE)
```

Creating Conservation Features Stack
```{r}
### reading in all habitat, geomorphological and species rasters

#####################
#######habitat#######
#####################
mangroves_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_mangrove.tif'))
seagrasses_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_seagrasses.tif'))
saltmarsh_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_saltmarsh.tif'))
coldcoral_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_coldcoral.tif'))
coral_TNC_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_coral_TNC.tif'))
mansalt_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_mansalt.tif'))
man_WAM_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_man_WAM.tif'))

habitat_stack <- stack(mangroves_rast, seagrasses_rast, saltmarsh_rast, coldcoral_rast, coral_TNC_rast, mansalt_rast, man_WAM_rast)

plot(habitat_stack)

### write raster
writeRaster(habitat_stack, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_habitat_stack.tif'), overwrite = TRUE)



############################
#######Turtle data#######
############################
turtle_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_turtle_all.tif'))
nest_may_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_nest_may.tif'))
nest_april_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_nest_april.tif'))
nest_march_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_nest_march.tif'))

turtle_stack <- stack(turtle_rast, nest_may_rast, nest_april_rast, nest_march_rast)

plot(turtle_stack)

### write raster
writeRaster(turtle_stack, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_turtle_stack.tif'), overwrite = TRUE)



##########################################
#######geomorphological features#######
##########################################
abyssal_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_abyssal.tif'))
insular_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_insular_shelf.tif'))
insular_upper_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_insular_upper.tif'))
isobata_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_isobata.tif'))
geomorph_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_geomorph_features.tif'))
plateaus_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_plateaus.tif'))
ridges_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_ridges.tif'))
spreading_ridges_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_spreading_ridges.tif'))

#KGA_features_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_KGA_features.tif'))
geomorph_stack <- stack(abyssal_rast, isobata_rast, plateaus_rast, ridges_rast, spreading_ridges_rast)

plot(geomorph_stack)

### write raster
writeRaster(geomorph_stack, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_geomorph_stack.tif'), overwrite = TRUE)


```

Creating Final Conservation Feature 
```{r}
### stack all species and habitat
features_stack <- stack(habitat_stack,turtle_stack, geomorph_stack)

### change into a dataframe to check it
features_df <- as.data.frame(features_stack)

### write raster
writeRaster(features_stack, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_features_stack.tif'),options="INTERLEAVE=BAND", overwrite = TRUE)
```

### Additional Files (Locked-in & Locked-out)
```{r}
### read in existing protected areas raster
pa_rast <- raster(here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/GA_pa.tif'))

plot(pa_rast)

```


### OBJECTIVE: Minimize area while meeting X% representation targets, locking out PAs and a boundary penalty of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective

## SCENARIO 1: Baseline - 10% target, Boundary Penalty = 0 
```{r}

####### habitat 10% ####### 
area_habitat_10 <- problem(GA_area, features = habitat_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_out_constraints(pa_rast) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve problem
sprob_area_habitat_10 <- solve(area_habitat_10)

## Plot the solution to see what it looks like
plot(sprob_area_habitat_10,  main = c("Area- 10% Habitat Targets"))

writeRaster(sprob_area_habitat_10, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/final_rasters/habitat_area_10.tif'), overwrite = TRUE)


####### turtles 10% ####### 
area_turtles_10 <- problem(GA_area, features = turtle_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(pa_rast) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve problem
sprob_area_turtles_10 <- solve(area_turtles_10)

## Plot the solution to see what it looks like
plot(sprob_area_turtles_10,  main = c("Area - 10% Turtles Targets"))

writeRaster(sprob_area_turtles_10, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/final_rasters/turtles_area_10.tif'), overwrite = TRUE)

####### geomorphological 20% ####### 
area_geomorph_10 <- problem(GA_area, features = geomorph_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(pa_rast) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve problem
sprob_area_geomorph_10 <- solve(area_geomorph_10)

## Plot the solution to see what it looks like
plot(sprob_area_geomorph_10,  main = c("Area- 10% Geomorphological Targets"))

writeRaster(sprob_area_geomorph_10, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/final_rasters/geomorph_area_10.tif'), overwrite = TRUE)

####### All features 10% ####### 
area_all_10 <- problem(GA_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(pa_rast) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve problem
sprob_area_all_10 <- solve(area_all_10)

## Plot the solution to see what it looks like
plot(sprob_area_all_10,  main = c("Area- 10% All Features Targets"))

writeRaster(sprob_area_all_10, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/final_rasters/ALL_area_10.tif'), overwrite = TRUE)

```


## SCENARIO 2: Baseline - 20% target, Boundary Penalty = 0 
```{r}

####### habitat 20% ####### 
area_habitat_20 <- problem(GA_area, features = habitat_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_out_constraints(pa_rast) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve problem
sprob_area_habitat_20 <- solve(area_habitat_20)

## Plot the solution to see what it looks like
plot(sprob_area_habitat_20,  main = c("Area- 20% Habitat Targets"))

writeRaster(sprob_area_habitat_20, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/final_rasters/habitat_area_20.tif'), overwrite = TRUE)

####### turtles 20% ####### 
area_turtles_20 <- problem(GA_area, features = turtle_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(pa_rast) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve problem
sprob_area_turtles_20 <- solve(area_turtles_20)

## Plot the solution to see what it looks like
plot(sprob_area_turtles_20,  main = c("Area- 20% Turtles Targets"))

writeRaster(sprob_area_turtles_20, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/final_rasters/turtles_area_20.tif'), overwrite = TRUE)

####### geomorphological 20% ####### 
area_geomorph_20 <- problem(GA_area, features = geomorph_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(pa_rast) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve problem
sprob_area_geomorph_20 <- solve(area_geomorph_20)

## Plot the solution to see what it looks like
plot(sprob_area_geomorph_20,  main = c("Area- 20% Geomorphological Targets"))

writeRaster(sprob_area_geomorph_20, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/final_rasters/geomorph_area_20.tif'), overwrite = TRUE)


####### All features 20% ####### 
area_all_20 <- problem(GA_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(pa_rast) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve problem
sprob_area_all_20 <- solve(area_all_20)

## Plot the solution to see what it looks like
plot(sprob_area_all_20,  main = c("Area- 20% All Features Targets"))

writeRaster(sprob_area_all_20, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/final_rasters/ALL_area_20.tif'), overwrite = TRUE)
```

## SCENARIO 3: Baseline - 30% target, Boundary Penalty = 0 
```{r}
####### habitat 30% ####### 
area_habitat_30 <- problem(GA_area, features = habitat_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.3) %>%
  add_locked_out_constraints(pa_rast) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve problem
sprob_area_habitat_30 <- solve(area_habitat_30)

## Plot the solution to see what it looks like
plot(sprob_area_habitat_30,  main = c("Area- 30% Habitat Targets"))
plot(sprob_area_habitat_20,  main = c("Area- 20% Habitat Targets"))
plot(sprob_area_habitat_10,  main = c("Area- 10% Habitat Targets"))

writeRaster(sprob_area_habitat_30, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/final_rasters/habitat_area_30.tif'), overwrite = TRUE)


####### turtles 30% ####### 
area_turtles_30 <- problem(GA_area, features = turtle_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.3) %>%
  add_locked_in_constraints(pa_rast) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve problem
sprob_area_turtles_30 <- solve(area_turtles_30)

## Plot the solution to see what it looks like
plot(sprob_area_turtles_30,  main = c("Area- 30% Turltes Targets"))

writeRaster(sprob_area_turtles_30, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/final_rasters/turtles_area_30.tif'), overwrite = TRUE)

####### geomorphological 30% ####### 
area_geomorph_30 <- problem(GA_area, features = geomorph_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.3) %>%
  add_locked_in_constraints(pa_rast) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve problem
sprob_area_geomorph_30 <- solve(area_geomorph_30)

## Plot the solution to see what it looks like
plot(sprob_area_geomorph_30,  main = c("Area- 30% Geomorphological Targets"))

writeRaster(sprob_area_geomorph_30, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/final_rasters/geomorph_area_30.tif'), overwrite = TRUE)


####### All features 30% ####### 
area_all_30 <- problem(GA_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.3) %>%
  add_locked_in_constraints(pa_rast) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve problem
sprob_area_all_30 <- solve(area_all_30)

## Plot the solution to see what it looks like
plot(sprob_area_all_30,  main = c("Area- 30% All Features Targets"))

writeRaster(sprob_area_all_30, here('C:/Users/vanessa_rathbone/Dropbox/Greater_Antilles/VR files/VR Files/outputs/final_rasters/all_area_30.tif'), overwrite = TRUE)

```


